---
title: "HW5"
author: "Melissa Luzardo"
date: "February 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#**MSDS 6306: Doing Data Science**
##Live Session Unit 05 Assignment

##1. Data Munging (30 points): Utilize yob2016.txt for this question. This file is a series of popular children's names born in the year 2016 in the United States. It consists of three columns with a first name, a gender, and the amount of children given that name. However, the data is raw and will need cleaning to make it tidy and usable.
##a. First, import the .txt file into R so you can process it. Keep in mind this is not a CSV file. You might have to open the file to see what you're dealing with. Assign the resulting data frame to an object, df, that consists of three columns with human-readable column names for each.

```{r yob2016A, include=TRUE}
#Set working directory
setwd("C:\\Users\\melis\\OneDrive\\Desktop\\SMU\\MSDS 6306 - Data Science\\Homework\\HW5")
#Import file and assing to "df"
df <- read.table('yob2016.txt', header= FALSE, sep= ";")
names(df) <- c("Name", "Gender", "Amount of Children")

```

##b. Display the summary and structure of df

```{r yob2016B, include=TRUE}
#Display summary and structure of df
summary(df)
str(df)
```

##c. Your client tells you that there is a problem with the raw file. One name was entered twice and misspelled. The client cannot remember which name it is; there are thousands he saw! But he did mention he accidentally put three y's at the end of the name. Write an R command to figure out which name it is and display it.

```{r yob2016C, include=TRUE}
#Find 3 y' at the end of name and difplay it.
grep(".yyy$", df[,1], value=TRUE)
```

##d. Upon finding the misspelled name, please remove this particular observation, as the client says it's redundant. Save the remaining dataset as an object: y2016

```{r y2016, include=TRUE}
#find the misspelled name, remove the row and assigned to y2016.
y2016 <- df[-which(df[,1]==grep(".yyy$", df[,1], value=TRUE)),]
```

##2. Data Merging (30 points): Utilize yob2015.txt for this question. This file is similar to yob2016, but contains names, gender, and total children given that name for the year 2015.
##a. Like 1a, please import the .txt file into R. Look at the file before you do. You might have to change some options to import it properly. Again, please give the dataframe human-readable column names. Assign the dataframe to y2015.

```{r y2015A, include=TRUE}
#Import file and assing to "y2015"
y2015 <- read.table('yob2015.txt', header= FALSE, sep= ",")
#Assign human-readable column names to dataframe
names(y2015) <- c("Name", "Gender", "Amount of Children")

```

##b. Display the last ten rows in the dataframe. Describe something you find interesting about these 10 rows.

```{r y2015B, include= TRUE}
#Display the last ten row
tail(y2015, 10)
```

Looks like the last ten names are all male and there are only 5 of each. That is interesting!!!

##c. Merge y2016 and y2015 by your Name column; assign it to final. The client only cares about names that have data for both 2016 and 2015; there should be no NA values in either of your number of children rows after merging.

```{r merge, include=TRUE}
#Merge y2016 and y2015 by Name
final <- merge(y2016, y2015, by=c("Name","Gender"))
#Rename columns for 2016 and 2015
names(final) <- c("Name", "Gender", "2016", "2015")
#Check for NA in y2016 and y2015 columns, remove if needed
sum(is.na(final[,3]))
sum(is.na(final[,4]))
```

##3. Data Summary (30 points): Utilize your data frame object final for this part.
##a. Create a new column called "Total" in final that adds the number of children in 2015 and 2016 together. In those two years combined, how many people were given popular names?

```{r Total, include=TRUE}
#Add 2016 and 2015 columns and assign to Total
Total <- final[,3]+final[,4]
#Adding Total to the final dataframe
final$Total <- Total
#Calculate to how many people were given popular name
sum(Total)
```

##b. Sort the data by Total. What are the top 10 most popular names?

```{r Sort, include=TRUE}
#Sort the data by Total
final<-final[order(Total, decreasing=TRUE), ] 
#Show top 10 most popular names
final$Name[1:10]
```

##c. The client is expecting a girl! Omit boys and give the top 10 most popular girl's names.

```{r Girl, include=TRUE}
#Ommit boys
final<-final[-which(final$Gender==grep("M", final$Gender, value=TRUE)), ]
```

##d. Write these top 10 girl names and their Totals to a CSV file. Leave out the other columns entirely.

```{r csv, include=TRUE}
#Create dataframe to for csv
Top10GirlNames <- data.frame(Name=final$Name[1:10],Total=final$Total[1:10])
#Write CSV file
write.csv(Top10GirlNames, file = "Top10GilrNames.csv", row.names = FALSE)
```

##4. Upload to GitHub (10 points): Push at minimum your RMarkdown for this homework assignment and a Codebook to one of your GitHub repositories (you might place this in a Homework repo like last week). The Codebook should contain a short definition of each object you create, and if creating multiple files, which file it is contained in. You are welcome and encouraged to add other files-just make sure you have a description and directions that are helpful for the grader.
##Github link https://github.com/meliluza/SMU-MSDS-6306/ HW5